### 2016-01-08 Paper/paper.tex
implemented suggested edits

%\documentclass[useAMS, usenatbib,usegraphicx,letter]{mn2e}
%\documentclass[11pt]{article}
\documentclass[reprint,aps,prd,superscriptaddress,showkeys,showpacs]{revtex4-1}
\usepackage{epsfig,amsmath,natbib}

\usepackage{aas_macros}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{dsfont}
\usepackage{hyperref}
\usepackage{color}
\usepackage{pbox}
\usepackage{booktabs}
\usepackage[dvipsnames]{xcolor}

\hypersetup{
	colorlinks=false,
	citecolor=green
}
% \usepackage{graphicx}
% \usepackage{epstopdf}
% \usepackage{natbib}

%%%%%%%%%%%%%%%%%
%Custom commands%
%%%%%%%%%%%%%%%%%

\newcommand{\bb}[1]{\mathbf{#1}}
\newcommand{\bbh}[1]{\mathbf{\hat{#1}}}
\newcommand{\h}[1]{\hat{#1}}
\newcommand\lsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
        \raise1pt\hbox{$<$}}}
\newcommand\gsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
        \raise1pt\hbox{$>$}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%ZH: I changed to more catchy title:
%\title{Simulating sample variance in weak lensing: effects on parameter inferences}
%AP: Ok

\title{Sample variance in weak lensing: how many simulations are required?}

\author{Andrea Petri}
\email{apetri@phys.columbia.edu}
\affiliation{Department of Physics, Columbia University, New York, NY 10027, USA}
\affiliation{Physics Department, Brookhaven National Laboratory, Upton, NY 11973, USA}

\author{Zolt\'an Haiman}
\affiliation{Department of Astronomy, Columbia University, New York, NY 10027, USA}

\author{Morgan May}
\affiliation{Physics Department, Brookhaven National Laboratory, Upton, NY 11973, USA}

\date{\today}

\label{firstpage}

\begin{abstract}

Constraining cosmology using weak gravitational lensing consists of
comparing a measured feature vector of dimension $N_b$ with its
simulated counterpart. An accurate estimate of the $N_b\times N_b$
feature covariance matrix $\bb{C}$ is essential to obtain accurate
parameter confidence intervals. When $\bb{C}$ is measured from a set
of simulations, an important question is how large this set should
be. To answer this question, we construct different ensembles of $N_r$
realizations of the shear field, using a common randomization
procedure that recycles the outputs from a smaller number $N_s\leq
N_r$ of independent ray-tracing $N$--body simulations.  We study
parameter confidence intervals as a function of ($N_s,N_r$) in the
range $1\leq N_s\leq 200$ and $1\leq N_r\leq 10^5$.  Previous work
\citep{DodelsonSchneider13} has shown that Gaussian noise in the
feature covariance lead, at quadratic order, to an $O(1/N_r)$
degradation of the parameter confidence intervals. Using a variety of
lensing features measured in our simulations, including shear-shear
power spectra and peak counts, we show that cubic and quartic
covariance fluctuations lead to additional $O(1/N_r^2)$ error
degradation that is not negligible when $N_r$ is only a factor of few
larger than $N_b$. We study the large $N_r$ limit, and find that a
single $N$--body simulation ($N_s=1$) can be repeatedly recycled to
produce as many as $N_r=10^5$ mutually independent shear maps. As a
result, a small number of simulations ($N_s=1$ or $2$) is sufficient
to forecast parameter confidence intervals at percent
accuracy.

%However, we find that in practice, a larger number of
%simulations might be needed to measure accurate feature means and to
%avoid parameter biases.
%ZH: I am not sure about the last sentence. I believe it is based on
%    an overly stringent reading of Fig.2, and that in fact we could
%    claim here, instead, that N_s=2 simulations are also sufficient
%    to get the correct mean.
%    See my comments at the end of the Conclusion section.

%AP: Ok, sentence removed

\end{abstract}

\keywords{Weak Gravitational Lensing --- Simulations --- Methods: numerical,statistical}
\pacs{98.80.-k, 95.36.+x, 95.30.Sf, 98.62.Sb}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%% INTRO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
%
Weak gravitational lensing (WL) is a promising cosmological probe for
constraining the dark energy equation of state $w$, and has been
considered by a range of past (CFHTLens \citep{cfht1,cfht2}, COSMOS \citep{cosmos}),
%ZH: also cite COSMOS survey here...
%AP: done
ongoing (DES \citep{DES}) and future (LSST \citep{LSST}, Euclid \citep{Euclid}, WFIRST \citep{WFIRST})
%ZH: also cite Euclid and WFIRST.
%AP: done
experiments. In an era where cosmology is data driven, accurate
numerical simulations of shear fields are becoming important for
several reasons, including the study of baryonic effects
\citep{BaryonXiuyuan,BaryonSemboloni,BaryonsWhite,BaryonsKnox,BaryonsZentner1,BaryonsZentner2}, 
%ZH: also cite Joop Schaye's hydro-simulation papers on lensing (Semboloni et al), and earlier papers by Martin White et al...  Lloyd Knox et al... and Zentner et al... (I think all of these are in Xiuyuan's paper).
%AP: done
the computation of non--Gaussian statistics
\citep{PeaksJan,MinkJan,MinkPetri,NG-Marian,NG-Jain1,NG-Jain2,NG-Jain3,NG-Refregier,NG-Dietrich} 
%ZH: Add here some citations of other groups - Marian+Bernstein.... Bhuvnesh Jain et al... Refregier et al... Dietrichh+Hartmann... Shirasaki+Yoshida.. (again just copy from our previous papers).
%AP: done
and the  assessment of various systematics
\citep{MinkShirasaki,Sys-Bard,Sys-Chang,Sys-Huterer}.
%ZH: cite papers by Jan+Morgan+Bard... and also Chihway Chan...Shirasaki+Yoshida.. and older ones by Dragan Huterer on photo-z errors..
%AP: done

A fundamental issue with predictions from simulations is that the
finite number of simulations naturally introduces fluctuations in the
forecasts, due to inevitable sample variance~\footnote{Sample variance is a broad term that has been used in the literature to describe a range of phenomena.  Throughout this paper, we use it to refer to the fluctuations in an ensemble of simulations, which represent random realizations of the same initial conditions.}.
%ZH: I think 'cosmic variance' usually refers to observations of the actual sky (which is finite). 
%AP: OK 
In general, quantities such as the mean or the variance of any feature
(e.g. the shear power spectrum at a multipole $\ell$), measured from a
finite set of simulations, will fluctuate, and can also suffer a bias.
While biases in the estimates of both the mean and the variance have
been studied extensively, the impact of fluctuations in the variance
has received comparatively less attention.  These fluctuations have,
however, been shown to have non-negligible effects on estimates of
features covariance matrices and hence on parameter constraints. In
particular, in the limit of Gaussian fluctuations, the parameter
confidence limits are degraged by a factor
$1+O(1/N_r)$~\citep{DodelsonSchneider13,Taylor12}.
%MM: also cite Taylor Joachimi and Kitching
%AP: done

This work studies these issues further, focusing on the number of
independent $N$--body simulations required for an accurate estimate of
the parameter constraints.  Ray-tracing simulations that resolve the
cosmic structures responsible for lensing on arcminute scales are
limited to physical sizes of hundreds of Mpc, and thus cover a solid
angle of only $O(10$ deg$^2)$.  As a result, many simulations are
required to tile a significant fraction of the sky, and to make
predictions for large ``all-sky'' surveys, such as the ones by DES,
LSST, Euclid, WFIRST.  In practice, this has led to the wide-spread
use of ``pseudo-independent'' realizations, i.e. a procedure in which
one randomizes and re-cycles the output of a single 3D simulation
multiple times.
%ZH: Add citations here of papers resorting to this technique
In light of the forthcoming large surveys, it is imperative to assess
the statistical validity of this approach, and to ask how many times a
single simulation can be fairly recycled.  In this paper, we address
these questions with ensembles of up to $N_r=10^5$ random
realizations, extracted from up to $N_s=200$ independent ray-tracing
N-body simulations. We focus in particular on the parameter $w$, and
on two different statistics: the (convergence) power spectrum and
the number counts of peaks.

This paper is organized as follows.  In \S~II, we summarize the shear
simulation methods we utilized, and describe the formalism we adopted
to forecast cosmological parameter constraints. We then vary the
number of simulations and the number of pseudo-independent
realizations, and present our main findings in \S~III. These results
are discussed further in \S~IV.  We offer our conclusions, and suggest
follow-up future work in \S~V.

%%%%%%%%%%%%%%%%%%% METHODS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Methods}

\subsection{Ray-tracing simulations of the convergence field}
\label{shearsim}
%
In this section, we describe how we constructed our shear field
ensembles. Background galaxies at redshift $z_s$ are lensed by large
scale structures between $z=0$ and $z_s$. The shape distortions due to
the cosmic shear $\pmb{\gamma}$ can be computed in terms of the dark
matter gravitational potential $\Phi(\bb{x},z)$. Because the evolution
of $\Phi$ with redshift is non--linear, it needs to be computed with
numerical simulations. We make use of the public code \texttt{Gadget2}
\citep{Gadget2}, with which we run a sequence of $200$ independent
dark--matter--only $N$--body simulations that track the evolution of
the density fluctuations. We assume a standard $\Lambda$CDM background
universe with the parameters
$(\Omega_m,\Omega_\Lambda,h,w,\sigma_8,n_s)=(0.26,0.74,0.72,-1,0.8,0.96)$.
We fix the comoving size of the simulation box to $240\mathrm{Mpc}/h$,
and use $512^3$ particles, corresponding to a Dark Matter particle mass of
$\approx 10^{10}M_\odot$.
%ZH: ``mass resolution'' here means DM particle mass ?
%AP: fixed

We assume a uniform galaxy distribution at a constant redshift $z_s=2$
(at which the simulation box has an angular size of $\theta_{\rm
  box}=3.5\mathrm{deg}$) and we discretize the mass distribution
between $z_s$ and the observer at $z=0$ with a sequence of 46 two
dimensional lenses of thickness $80\mathrm{Mpc}/h$. The surface
density on each lens plane is computed by projecting the
three--dimensional density measured from \texttt{Gadget2}
snapshots. We then apply the multi--lens--plane algorithm (see
\citep{RayTracingHartlap,RayTracingJain} for example) to trace the
deflections of $n_{\rm ray}^2=2048^2$ light rays arranged on a square grid
of total size $\theta_{\rm box}$, from $z=0$ to $z_s$. Our
implementation of this algorithm is part of the \texttt{LensTools}
computing package we have been developing \citep{LensTools}, and have
released under the \texttt{MIT} license. Many different realizations
$r$ of the same shear field $\pmb{\gamma}_r(\pmb{\theta})$ can be
generated by picking different lens planes that lie between the
observer and $z_s$. The randomization procedure we adopt is the
following (see \citep{Sato12} for reference):
%ZH: we should cite some paper here. We didn't invent most of this
%procedure ourselves (except for maybe the 3-times thicker simulation
%output, and the random mixing of the planes from different N-body sims ?).
%AP: I added the citation to Sato et. al, which refer to this procedure without citing anyone

\begin{itemize}
\item For each lens-plane redshift $z_l$, select the snapshot at $z_l$ from the $i$--th $N$--body simulation, where $i$ is a random integer $i\in [1,N_s]$.
\item Choose randomly between the three orthogonal directions ${\bb{n}_x,\bb{n}_y,\bb{n}_z}$: the lens plane will be perpendicular to this direction.
\item Choose the position of the plane along the snapshot: because the lens thickness is 1/3 the size of the box, we can cut three different slices of the simulation box for each orientation ${\bb{n}_x,\bb{n}_y,\bb{n}_z}$. This gives a total of 9 choices for generating a lens plane out of a single $N$--body snapshot.
\item Perform a periodic random shift of the lens plane along its two directions.
%ZH: should we specify the range of the random shifts (0 to 1/2 the lens plane dimension?)
%AP: This is not necessary, since the shifts are periodic, in principle one could choose a random number of pixels. In particular this number should be between 0 and 4095, because each lens plane has 4096 pixels
\item Repeat the above procedure for each lens-plane redshift $z_l$.
\end{itemize}  
%
This randomization procedure allows us to produce an (almost)
arbitrary number $N_r$ of shear realizations
$\pmb{\gamma}_r(\pmb{\theta})$. However, these realizations are not
guaranteed to be independent, if $N_s$ is not large enough. Using the
set of 200 independent $N$--body simulations, we construct different
ensembles with different choices of $N_s\in[1,200]$. Each of these
ensembles consists of the same number $N_r=1000$ of shear
realizations. We also build an additional ensemble with $N_s=1$ and
$N_r=10^5$ realizations. For each realization of each ensemble, we
reconstruct the convergence $\kappa_r(\pmb{\theta})$ from the trace of the light ray deflection Jacobian matrix, measured looking at the difference in angle deflection between nearby light rays \citep{RayTracingHartlap,RayTracingJain,Sato12}.
%ZH: we should add 2 sentences on how you measured kappa.  Directly from the ray-tracing?
%   (but above, when you discuss ray-tracing, you mention only gamma).  Or did you do 
%    Kaiser-Squires on the gamma field ? 
%AP: done, we measure kappa directly from the ray tracing

We measure the $\kappa$ angular power spectrum $P^{\kappa\kappa}_r(\ell)$ defined as
\begin{equation}
\langle\tilde{\kappa}_r(\pmb{\ell})\tilde{\kappa}_r(\pmb{\ell}')\rangle = (2\pi)^2\delta_D(\pmb{\ell}+\pmb{\ell}')P^{\kappa\kappa}_r(\ell)
\end{equation}
%
As an additional summary statistic, we consider the counts of local
$\kappa$ maxima of a certain height $\kappa_0$, $n_r(\kappa_0)$
(hereafter \textit{peak counts}), with varying $\kappa_0$ chosen between the minimum and maximum values measured from the maps $(\kappa_{\rm min},\kappa_{\rm max})=(-0.06,0.45)$. Different choices of $\kappa_0$ binning used in this work are outlined in Table \ref{featuretable}. 
%ZH: state here the range and binning choice of $\kappa_0$
%AP: done
The fact that the ensemble of $N_r$ realizations is not completely
independent if $N_s$ is not large enough can have an effect on the
covariance estimators of both $P^{\kappa\kappa}$ and $n(\kappa_0)$.

%ZH: it would be logical to mention the other w,Sig,Om, sims here, and point to \ref{simtable})
%AP: done, added the below

To measure the cosmological dependence of the $\kappa$ peak counts, we also performed additional ray--tracing simulations with different combinations of the cosmological parameter triplet $(\Omega_m,w,\sigma_8)$. A summary of the complete set of shear ensembles used in this work is outlined in Table \ref{simtable}.

\subsection{Cosmological parameter inference}
%
Let $\bbh{d}$ be a single estimate for a feature of dimension $N_b$, 
%ZH: ``Hat'' here refers to an estimator?
%AP: Yes: quantities without the hat are the expectation values 
$\bb{d}(\bb{p})$ be the
%smooth expectation
%ZH: I believe ``smooth expectation'' means true mean value, yes?
%AP: yes
true value of this feature at a point $\bb{p}$ in parameter space
(which has a dimension $N_p$) and $\bb{C}$ be the $N_b\times N_b$
feature covariance matrix. For the purpose of this work $\bb{p}$ is
the triplet $(\Omega_m,w,\sigma_8)$ and $\bb{d}$ is one of the
features -- either a power spectrum or a peak count histogram -- in Table
\ref{featuretable}. Although existing emulators (see
\citep{coyote2,Nicaea,SchneiderKnoxCovariance} for example) can be used to compute $\bb{d}(\bb{p}),\bb{C}$ in some cases, this is not true in the general case. 
%ZH: I am missing the point here.  Why is covariance matrix ``more complicated'' ?
%    Even to estimate a mean, one has to ``generate a series of mock realizations''.
%    Unless you are willing to equate a single realization as the mean.   But then
%    by the same logic, you could estimate the covariance matrix from two realizations.
%    Please clarify what you mean...  maybe I am missing something?  Or you simply mean
%    that published emulators usually do not emulate the cov.matrix (even though they could)?
%    Note that there is a paper by Schneider and Knox, in which they do an interpolation of the
%    power spectrum covariance matrix.
%AP: The intended meaning is the latter: published emulators only predict the means. It is also true that getting the covariance accurate is more difficult than getting the means accurate. I rephrased the sentence.
Estimating $\bb{C}$ from simulations involves generating a series of mock realizations $\bbh{d}_r$ with $r=1...N_r$ and computing the
sample covariance $\bbh{C}$,

\begin{equation}
\bb{\bar{d}} = \frac{1}{N_r}\sum_{r=1}^{N_r} \bbh{d}_r,
\end{equation}

\begin{equation}
\label{covest}
\bbh{C} = \frac{1}{N_r-1}\sum_{r=1}^{N_r} (\bbh{d}_r - \bar{\bb{d}}) (\bbh{d}_r - \bar{\bb{d}})^T.
\end{equation}
%
Assuming a normal feature likelihood, together with a flat prior on
the parameter space, the parameter posterior distribution
$\mathcal{L}(\bb{p}\vert\bbh{d}_{\rm obs})$ given an observed instance of
$\bbh{d}$, which we call $\bbh{d}_{\rm obs}$, follows from Bayes' theorem,
%
\begin{equation}
\label{posteriorbayes}
-2\log\mathcal{L}(\bb{p}\vert\bbh{d}_{\rm obs}) = [\bbh{d}_{\rm obs}-\bb{d}(\bb{p})]^T\bbh{C}^{-1}[\bbh{d}_{\rm obs}-\bb{d}(\bb{p})].
\end{equation}
%
For the sake of simplicity, we approximate the posterior as a Gaussian
around its maximum. This corresponds to Taylor-expanding the simulated
feature to first order around a point $\bb{p}_0$ (ideally the maximum
of Eq.~\ref{posteriorbayes}):
%
\begin{equation}
\bb{d}(\bb{p}) \approx \bb{d}_0 + \bb{d}_0^\prime(\bb{p}-\bb{p}_0).
\end{equation}
%

To measure the derivatives of the features $\bb{d}'_0$ with respect to
the cosmological parameters, we make use of the public code
\texttt{Nicaea} \citep{Nicaea} for the power spectrum, and we use an
independent simulation set (containing simulations with a variety of
different combinations of $(\Omega_m,w,\sigma_8)$, see Table
\ref{simtable}) for the peak counts.

\begin{table*}
\begin{center}
\begin{tabular}{c|c|c||c|c}
\toprule
$\Omega_m$ &  $w$ & $\sigma_8$ & $(N_s,N_r)$ & Number of $\kappa$ ensembles \\ \hline \hline
\midrule
%
0.26 & $-$1 & 0.8 & $(1\div200,1024),(1,128000)$ & 16 \\
0.29 & $-$1 & 0.8 & (1,1024) & 1 \\
0.26 & $-$0.8 & 0.8 & (1,1024) & 1 \\
0.26 & $-$1 & 0.6 & (1,1024) & 1 \\
%
\bottomrule
\end{tabular}
\end{center}
\caption{Summary of the shear ensembles used in this work.}
%ZH: So each of these have N_r=1000, yes?  You should state this here. And also,
%    Add the ensemble with N_s=1, N_r=10^5 to this table.
%    Also: point the reader to this table in the text, up front, when the 
%    ensembles are first discussed.
%AP: done
\label{simtable}
\end{table*}

We can build the estimator for the posterior maximum $\bbh{p}$, given
the observation $\bbh{d}_{\rm obs}$, as follows:
%
\begin{equation}
\label{estimatormean}
\bbh{p} = \bb{p}_0 + \bbh{T}(\bbh{d}_{\rm obs}-\bb{d}_0),
\end{equation}

\begin{equation}
\bbh{T} = (\bb{d}_0'^T\bbh{C}^{-1}\bb{d}_0')^{-1}\bb{d}_0'^T\bbh{C}^{-1}.
\end{equation}
%
Because $\bbh{p}$ is estimated using a single noisy data instance
$\bbh{d}_{\rm obs}$, its estimate will be scattered around the true
value $\langle\bbh{p}\rangle_O$. In the following we use the
$\langle\rangle_O$ notation for expectation values taken with respect
to observations, while we keep the notation $\langle\rangle$ for
expectation values taken with respect to the simulations. Defining the
\textit{precision matrix} $\bbh{\Psi}=\bbh{C}^{-1}$, we can express
the estimator of the observational scatter in $\bbh{p}$:

\begin{equation}
\label{estimatorcovariance}
\h{\Sigma}_\bb{p} = \bbh{F}^{-1}\bb{d}_0'^T\bbh{\Psi}\langle(\bbh{d}_{\rm obs}-\bb{d}_0)(\bbh{d}_{\rm obs}-\bb{d}_0)^T\rangle_O\bbh{\Psi}\bb{d}_0'\bbh{F}^{-1},
\end{equation}

\begin{equation}
\label{estimatorfisher}
\bbh{F} = \bb{d}_0'^T\bbh{\Psi}\bb{d}_0'.
\end{equation}
%
Here we introduced the familiar Fisher matrix estimator $\bbh{F} =
\bb{d}_0'^T\bbh{\Psi}\bb{d}_0'$ and, for simplicity, we assumed
$\langle\bbh{d}_{\rm obs}\rangle_O=\bb{d}_0$, so that
$\langle(\bbh{d}_{\rm obs}-\bb{d}_0)(\bbh{d}_{\rm
  obs}-\bb{d}_0)^T\rangle_O=\bb{C}$.  When we perform an observation
$\bbh{d}_{\rm obs}$, the parameter estimate $\bbh{p}$ is a random draw
from a probability distribution with variance $\h{\Sigma}_\bb{p}$,
which inherits noise from the simulations. The noise in the covariance
estimator (Eq.~\ref{covest}) and in its inverse $\bbh{\Psi}$ propagate
all the way to the posterior (Eq.~\ref{posteriorbayes}), the parameter
estimate (Eq.~\ref{estimatormean}) and its variance
(Eq.~\ref{estimatorcovariance}). Following
refs.~\citep{DodelsonSchneider13,Taylor12,MasumotoWishart} we compute
the expectation value of Eq.~(\ref{estimatorcovariance}) over
simulations, $\langle\h{\Sigma}_\bb{p}\rangle$, up to $O(1/N_r^2)$ by
expanding Eq.~(\ref{estimatorcovariance}) to quartic order in the
statistical fluctuations of $\bbh{\Psi}$. Denoting the noiseless
parameter covariance, i.e. the usual Fisher matrix, as
$\Sigma_\bb{p}=\bb{F}^{-1}$, we find the result~\footnote{The details of the calculation are given in Appendix A.}:

\begin{widetext}
\begin{equation}
\label{quarticdegradation}
\langle\h{\Sigma}_\bb{p}\rangle = \Sigma_\bb{p}\left[1+\frac{N_b-N_p}{N_r}+\frac{(N_b-N_p)(N_b-N_p+2)}{N_r^2}\right] + O\left(\frac{1}{N_r^3}\right).
\end{equation}
\end{widetext}
%
Next, we restrict ourselves to the large $N_r$ limit, and we further
investigate the behavior of the $O(1/N_r)$ term. We consider three
cases \footnote{The details of these calculations are given in Appendix B}:

\begin{enumerate}
\item If the true data covariance $\bb{C}$ is known, the estimator in eq.~(\ref{estimatorcovariance}) is biased, and the dominant contribution of the bias comes from the second order fluctuations in $\bbh{\Psi}$. Once the expectation values over simulations are taken, the bias sums up to

\begin{equation}
\label{dodelsonscaling}
\langle\h{\Sigma}_\bb{p}\rangle=\Sigma_\bb{p}\left(1+\frac{N_b-N_p}{N_r}\right).
\end{equation}
%
This is the result obtained by ref.~\citep{DodelsonSchneider13}.

\item Usually the true data covariance is unknown, and it is tempting to plug in its estimator $\bbh{C}$, measured from the same simulation set we use to compute $\bbh{\Psi}$. This approach has been used before in the literature (e.g.~\citep{MinkPetri,MinkShirasaki}). If this is done without correcting for the bias in $\bbh{\Psi}$ (see eq.~(\ref{firstmoment}) below or \citep{Taylor12} for reference), 
%ZH: we should probably cite Hartlap et al. and Taylor et al. here for this Psi bias.  (I assume this is what you mean in eq. 19)
%AP: done
the parameter variance will have a contribution from both the second and first-order fluctuations in $\bbh{\Psi}$, which now have a nonzero expectation value. In this case the bias sums up to
%
\begin{equation}
\label{mockscalinguncorrected}
\langle\h{\Sigma}_\bb{p}\rangle=\Sigma_\bb{p}\left(1-\frac{N_b-N_p}{N_r}\right).
\end{equation}

\item If we repeat the same exercise as above, but we correct for the bias in the precision matrix estimator, we are left with

\begin{equation}
\label{mockscalingcorrected}
\langle\h{\Sigma}_\bb{p}\rangle=\Sigma_\bb{p}\left(1+\frac{1+N_p}{N_r}\right).
\end{equation}

\end{enumerate} 
%
The error degradation in each parameter $p$, at leading order, scales
as $D/N_r$, where $D=N_b-N_p$, $N_p-N_b$, and $1+N_p$ for cases 1, 2,
and 3, respectively.
%ZH: It was not clear to me what the main point of the text below was. I paraphrased it slightly. OK?
%This means that, when parameter errors are
%estimated purely from simulations, and the same simulation set is used
%to estimate both $\bbh{C},\bbh{\Psi}$, these errors tend to decrease
%with $N_b$ (case 2), or at best remain constant (case 3), while the
%distribution of $\bbh{p}$ has still a variance that increases with
%$N_b$. This typically leads to underestimates in parameter errors. 
%AP: Not exactly: my point was that the distribution of \bbh{p} has a true variance that increases with Nb, but the estimate of this variance (which we quote as the errorbar) does not increase with Nb. Hence we get the wrong parameter fit (i.e. which excludes the true value at high significance level)
Note that in the last case, which is most relevant when fitting actual data, the estimated degradation turns out to be too optimistic: the parameter estimate $\bbh{p}$ has a variance that grows linearly with $N_b$, but the estimated degradation in eq. (\ref{mockscalingcorrected}) is constant with $N_b$. This can lead to underestimation of error bars, which can be mistakenly interpreted as a parameter bias.  
We derived the scaling relations of the form
%
\begin{equation}
\label{ourscaling}
\langle\h{\sigma}_p^2\rangle = \sigma^2_{p,\infty}(N_s)\left(1+\frac{D}{N_r}\right)
\end{equation}
%
above by assuming that the noise in $\bbh{d}$ is Gaussian. We then directly test these scaling relations using our simulations, in the limits of both high and low $N_r$.
%ZH: It is still a little unclear here what is being tested.  Is it the Gaussianity assumption? I added that.
%    But perhaps you had some other issue in mind.
%AP: The noise in $\bbh{C}$ and $\bbh{\Psi}$ is not Gaussian: if the data vector from which $\bbh{C}$ and $\bbh{\Psi}$ are estimated in Gaussian, then these matrices are distibuted according to, respectively, the Wishart and inverse Wishart distributions. We are testing that the degradation effectively decreases as 1/Nr at leading order.  
We indicate the diagonal elements of $\h{\Sigma}_{\bb{p}}$ as $\h{\sigma}^2_p=\mathrm{diag}(\h{\Sigma}_{\bb{p}})$ 
%ZH: I am confused.  Why is a new symbol \sigma necessary here, on top of \Sigma ?
%    (And also: here you write \Sigma_pp, but above it was denoted \Sigma_p  )
%AP: fixed. I introduced the \sigma symbol for the single parameter variances (as opposed to the full parameter covariance), because these are the quantities we plot
and we indicate by $\sigma^2_{p,\infty}$ the expectation value of the variance of each parameter in the limit of an infinite number of realizations
$N_r\rightarrow\infty$.  We call $D$ the \textit{effective dimensionality} of the feature space (which, as seen before, can be
negative in some pathological cases).
%ZH: I added pathological here.  Do you agree?  It's only negative in case 1, when
%    N_p > N_b (which is impossible -- cannot measure N_p parameters with N_b observables)
%    or in case 2 (when N_b > N_p, so that's OK, but this case ust means we ``forgot'' to correct for the psi-bias).
%AP: agreed
We compute the expectation values of $\h{\sigma}^2_p$
(eqs.~\ref{estimatorcovariance} and \ref{ourscaling}) by
averaging over 100 random resamplings of our shear ensembles. For the
true feature covariance matrix $\langle(\bbh{d}-\bb{d}_0)(\bbh{d}-\bb{d}_0)^T\rangle=\bb{C}$ we use
the estimated covariance from a grand ensemble built with the union of
all the ensembles with different $N_s$.

The true parameter variance $\sigma^2_{p,\infty}(N_s)$ in principle
can depend on the number of independent $N$--body simulations $N_s$,
which appears in the randomization procedure described in
\S~\ref{shearsim} above. This is because if $N_s$ is not large enough,
the different shear realizations cannot be all independent, and hence
the true variance $\sigma^2_{p,\infty}(N_s\rightarrow\infty)$ cannot
be recovered for low $N_s$ even if $N_r$ is arbitrarily large. In the
next section, we present our main findings.

%%%%%%%%%%%%%%%%%%% RESULTS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure*}
\includegraphics[scale=0.4]{Figures/ps_pdf.eps}
\caption{PDF of the $\kappa$ power spectrum
  $\mathcal{L}(P_l^{\kappa\kappa})$ at four selected multipoles
  $\ell=115,344,1349,5283$, for different shear ensembles constructed
  from on $N_s=$1 (black), 2 (blue), 5 (green), 50 (red), and 100
  (purple) independent N-body simulations. Each curve is based on
  $N_r=1024$ realizations.
%ZH: I added the Nr=1000 above. Yes?
%AP: Yes, 1024 in reality, but it doesn't change much
  The dashed black curves correspond to ensembles generated with $N_s=1$ and
  $N_r=128000$. For $N_s\geq2$, the distributions appear
  similar to the eye; this similarity is confirmed by the comparisons
  in Figures \ref{means_ns} and \ref{ps_var} below.}
%MM: added last two sentences
%MM: Thick black line draws the eye and looks like the most significant thing on the plot. replace with dotted black line or something else that stands out less.
%ZH: Also, include a label ``N_r=10^5'' with the other N_s=... labels in the figure panels 
%   (it is just confusing why there are labels for all the other curves, but not for the thick black)
%AP: done
\label{ps_pdf}
\end{figure*}

\begin{figure}
\includegraphics[scale=0.3]{Figures/means_ns.eps}
\caption{The mean value $\bb{\bar{d}}$
% $d$  
%ZH: I think you mean  \bb{\bar{d}} here?  You introduced a lot of different
%    subtle notations in the previous section, so we should  follow those ....
%AP: OK
  of various features, measured from ensembles created from different
  numbers $N_s$ of simulations.  For each case, the difference compared
  to the mean in the $N_s=200$ ensemble is shown, in units of the
  statistical error measured in the $N_s=200$ ensemble.  The colored
  curves refer to shear--shear power spectra measured at $\ell=115$
  (black), 1027 (cyan), and 5283 (green), and peak counts 
%ZH: I replace ``of'' with ``above''.  These are cumulative peak counts N(>kappa_0) ?
%    Or do you mean differential counts in some delta-kappa_0 bins?   Then you should
%    correct the text (and state the bin width).  Table II contains some kappa-bins
%    but no obvious relation to the 0.05, 0.17  and 0.28 mentioned here...
%AP: The peak counts are differential, I added the clarifications; I chose three values of kappa0 for the sake of presentation simplicity (one value in the low height peaks, one in the middle and one in the large)
  with height of $\kappa_0=0.05$ (red), 0.17 (purple), and 0.28
  (orange). The $\kappa$ bin width for the peak counts has been fixed to $\Delta\kappa=0.011$. The dashed black line shows a level of $0.1\sigma$ accuracy for reference.  For $N_s\geq 2$, the means are statistically indistinguishable (at $\sim0.1\sigma$) from the means measured in the ensemble with $N_s=200$.}
%MM: added last sentence stating conclusion to be drawn from the figure
%ZH:  Good idea.  All figures captions should similarly state a conclusion!
%ZH:  What is the horizontal dashed line?  It is never mentioned in the caption or
%     the text.   I guess you decided to mark the ``0.1 sigma'' deviation?
%     We should state that.    I suppose this is what Morgan refers to by 
%    ``statistically indistinguishable'' ....  0.1 sigma.
%AP: Ok, added the clarifications to the caption
\label{means_ns}
\end{figure}

\begin{figure}
\includegraphics[scale=0.3]{Figures/ps_variance.eps}
\caption{Variance of the $\kappa$ power spectrum as a function of the
  multipole $\ell$, in units of the expected Gaussian variance from
  equation (\ref{gaussianvar}). The variance is measured from
  different shear ensembles based on $N_s=$1 (black), 2 (blue), 5
  (green), 10 (red), 50 (purple), or 100 (orange) N--body simulations. 
%ZH: Added ``conclusion'':
%AP: OK
  Non--Gaussianities of the underlying structures increase the variance on
  small scales, but no clear trend with $N_s$ can be identied on any scale.}
\label{ps_var}
\end{figure}

\begin{figure}
\includegraphics[scale=0.3]{Figures/curving_nb.eps}
\caption{Expectation value of the variance of $w$ computed from
  equation (\ref{estimatorcovariance}), shown as a function of
  $1/N_r$. The different symbols and colors correspond to the features
  listed in Table~\ref{featuretable}.
%ZH: It would be better to write the features on the legend embedded in the figure.
%    Why make the reader flip back-and-forth between the figure and the Table ?
%AP: done
  The dashed and solid curves show the analytic predictions from
  equation (\ref{quarticdegradation}) at orders $O(1/N_r)$ and
  $O(1/N_r^2)$, respectively. The asymptotic variance
  $\sigma^2_{w,\infty}$ has been computed from a linear regression of
  $\langle\h{\sigma}^2_w\rangle$ vs $1/N_r$ for $N_r>500$. The figure
  clearly shows that terms beyond $O(1/N_r)$ need to be considered,
  unless $N_r\gg N_b$.}
%MM: added last sentence stating conclusion to be drawn from the figure
%in legend box change line that reads Nb=7 divided by 8
%AP: done, the legend is comprehensive now
\label{curvingnb}
\end{figure}

\begin{figure}
\includegraphics[scale=0.3]{Figures/scaling_nr.eps}
\caption{Bias in the variance of $w$,
  $\langle\h{\sigma}^2_w\rangle-\sigma^2_{w,\infty}$, as a function of
  the number of realizations $N_r$ used to estimate the covariance
  (eq.~\ref{covest}). Results for different ensembles, generated from
  $N_s=$1 (black), 2 (blue), 5 (green), 50 (red), and 100 (purple)
  simulations are shown. The figure shows both the measured trends
  (solid lines) and their best fits based on eq.~(\ref{ourscaling}),
  with $D$ treated as a free parameter.
%ZH: We need to explain what you mean by best-fit (presumably you treat ``D'' as a free parameter).
%   I added that. But if you meant something else, please correct.
%AP: OK
  The thick black curve corresponds to an ensemble of $N_r=128000$
  realizations from a single simulation ($N_s=1$).
%ZH: I added 'conclusion':
%AP: OK
  No systematic deviation from the expected scaling is seen, even in
  the extreme last case.  }
\label{wvar_nr}
\end{figure}

\begin{figure}
\includegraphics[scale=0.3]{Figures/scaling_ns.eps}
\caption{The variance of $w$ in the limit of $N_r\rightarrow\infty$ (measured from the intercept of the fit $\sigma^2_w$ vs $1/N_r$),
%ZH: state here what N_r was actually used to approximate the infinity limit
%AP: the value plotted is actually $N_r\rightarrow\infty$, because it is estimated from the intercept of the fit. Even if all the points have Nr finite, the intercept of the line corresponds to Nr infinite. 
  varying the number of simulations $N_s$ used in the ensemble to
  estimate the covariance (eq.~\ref{covest}). We show the dependence
  of $\sigma_{w,\infty}^2(N_s)$ in units of the mean over the 16 ensembles with different $N_s$, 
%ZH: Just to make sure: what you mean here is an averaging of sigma_infty over the 16 different
%    cases of N_s ?   
%AP: Yes
  for the power spectrum logarithmically binned (black, $N_b=15,\ell\in[100,6000]$), the power spectrum linearly binned (red,$N_b=39,\ell\in[100,6000]$) and the peak counts (green,$N_b=45,\kappa_0\in[-0.06,0.45]$).
%ZH: What ell and kappa ranges were used?
%ZH: Added conclusion:
%AP: OK, added binning information
  No trend with $N_s$ is seen for $N_s\geq 2$, and the differences are only of order 1\%.}
\label{wvar_ns}
\end{figure}

\begin{table*}
\begin{center}
\begin{tabular}{ccccc}
\toprule
\textbf{Feature} &  \textbf{Specifications} & $N_b$ &  \textbf{Symbol} & \textbf{Color} \\ \hline \hline
\midrule
Power Spectrum, log binning  & $\ell \in [100,800] $ & 8 & $\times$ & black  \\ 
Power Spectrum, log binning  & $\ell \in [1000,6000] $ & 7 & $\blacksquare$ & black  \\ 
Power Spectrum, log binning  & $\ell \in [100,6000] $ & 15 & \textcolor{red}{$\bullet$} & \textcolor{red}{red}  \\
Power Spectrum, linear binning  & $\ell \in [100,2000] $ & 15 & \textcolor{red}{$+$} & \textcolor{red}{red}  \\ 
Power Spectrum, linear binning  & $\ell \in [2500,4500] $ & 15 & \textcolor{red}{$\times$} & \textcolor{red}{red}  \\
Power Spectrum, linear binning  & $\ell \in [100,4500] $ & 30 & \textcolor{OliveGreen}{$\bullet$} & \textcolor{OliveGreen}{green}  \\ 
Power Spectrum, linear binning  & $\ell \in [100,6000] $ & 39 & \textcolor{blue}{$\bullet$} & \textcolor{blue}{blue}  \\ \hline
Low peaks  & $\kappa_0 \in [-0.06,0.09] $ & 15 & \textcolor{red}{$+$} & \textcolor{red}{red}  \\ 
Intermediate peaks  & $\kappa_0 \in [0.1,0.27] $ & 15 & \textcolor{red}{$\bigstar$} & \textcolor{red}{red}  \\ 
High peaks  & $\kappa_0 \in [0.28,0.45] $ & 15 & \textcolor{red}{$\diamond$} & \textcolor{red}{red}  \\
Low+Intermediate peaks  & $\kappa_0 \in [-0.06,0.27] $ & 30 & \textcolor{OliveGreen}{$\times$} & \textcolor{OliveGreen}{green}  \\
Intermediate+High peaks  & $\kappa_0 \in [0.1,0.45] $ & 30 & \textcolor{OliveGreen}{$\blacksquare$} & \textcolor{OliveGreen}{green}  \\
All peaks  & $\kappa_0 \in [-0.06,0.45] $ & 45 & \textcolor{magenta}{$\blacksquare$} & \textcolor{magenta}{magenta}  \\ \hline
\bottomrule
\end{tabular}
\end{center}
\caption{Catalog of feature types used in this work, along with the chosen number of bands $N_b$ and the plot legends for Figure \ref{curvingnb}.}
\label{featuretable}
\end{table*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this section we present the main results of this work. We show the
qualitative behavior of a variety of feature $\bbh{d}_r$ probability
distribution functions (PDFs) in ensembles built with different $N_s$
and $N_r$. In Figure~\ref{ps_pdf}, we show the PDF of the power
spectrum at four selected multipoles, spanning the linear ($\ell=115$)
to the nonlinear ($\ell=5283$) regime. In Figure~\ref{means_ns}, we
shows the ensemble mean for these power spectra, as well as for peak
counts of three different $\kappa_0$ heights (corresponding to
$\approx 2\div13\sigma$ peaks), as a function of $N_s$.
%ZH: added above 2-12.  OK ?
%AP: the actual numbers are 2-13, but again, only minor change
In Figure~\ref{ps_var}, we show the variance of the power spectrum at
each multipole, as a function of $N_s$, in units of the variance
expected if the convergence $\kappa$ was a Gaussian random field
%
%ZH:  I changed ``l''  to ``\ell'' in many places, for the multipole
%AP: OK, I fixed this throughout the rest of the paper
\begin{equation}
\label{gaussianvar}
\mathrm{Var}(P^{\kappa\kappa}_\ell) = \frac{(P^{\kappa\kappa}_\ell)^2}{N_{\rm eff}(\ell).}
\end{equation}
%
Here $N_{\rm eff}(l)$ is the number of independent modes used to
estimate the power spectrum at $\ell$.

In practice, we measure $P^{\kappa\kappa}_\ell$ on the Fourier
transform of the pixelized simulated map $\kappa_r(\pmb{\theta})$,
using the FFT algorithm, and some care must be taken to count the
number of modes $N_{\rm eff}(\ell)$ correctly. Each pixel $(i_x,j_y)$
in Fourier space corresponds to a mode
$(\ell_x,\ell_y)=2\pi(i_x,i_y)/\theta_{\rm box}$, with $i_x=-n_{\rm
  ray}/2,...,n_{\rm ray}/2$ and $i_y=0,...,n_{\rm ray}/2$.
%ZH: added this:
%AP: OK
Here $n_{\rm ray}=2048$ is the linear number of pixels on the
ray--traced convergence maps.  We count the number of pixels $N(\ell)$
that fall inside a multipole bin $(\ell_1,\ell_2)$. Because the
$\kappa$ field is real, the modes $(\pm \ell_x,0)$ are not
independent. If we let $N(\ell,\ell_y=0)$ be the number of
non--independent modes, the effective number of independent modes
for the variance is given by

\begin{equation}
N_{\rm eff}(\ell) = \frac{N^2(\ell)}{N(\ell)+N(\ell,\ell_y=0)}.
\end{equation}
%
This correction is important at low $\ell$, where pixelization effects
are non-negligible; $N_{\rm eff}(\ell\gg2\pi/\theta_{\rm box})\approx
N(\ell)$.

In Figures~\ref{curvingnb} and \ref{wvar_nr}, we show the dependence
of the confidence range $\langle\h{\sigma}^2_w\rangle$ on $N_r$,
derived from the features used in this work (see Table \ref{featuretable} for a comprehensive list). Figure \ref{wvar_nr} shows the $N_r$ trends of the $w$ constraint obtained with the power spectrum logarithmically binned with $\ell\in[100,6000],N_b=15$. 
%ZH: previously you only stated ``power spectrum'' for Figure 5.  But I assume Fig.4 is
%    also based on the power spectrum?  If not, please correct.
%AP: OK, fixed
Figure~\ref{curvingnb} illustrates the behavior at relatively low
$N_r$, and compares $\langle\h{\sigma}^2_w\rangle$ measured directly
from the simulations with the analytic expectations from
equation~(\ref{quarticdegradation}).  Figure~\ref{curvingnb} shows the
behavior in the limit of a large number $N_r\gg500$ of realizations,
and compares it with the scaling of the form in
equation~(\ref{ourscaling}). Finally, in Figure~\ref{wvar_ns}, we show
how the $w$ confidence limit changes with $N_s$.

%%%%%%%%%%%%%%%%%% DISCUSSION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Discussion}

In this section we discuss our main findings and their
implications. Figure \ref{ps_pdf} shows that, although different
choices of $N_s$ do not affect the power spectrum PDF on large scales
(top two panels), there are some qualitative differences on smaller
scales (bottom two panels). On these smaller scales, shear ensembles
built from $N_s=1$ do not produce the same statistical behavior as
ensembles built with larger $N_s$. In particular, looking at the
black curves, we see that the $N_s=1$ ensembles exibit large shifts
with respect to the other PDFs to lower power, including the locations
of the peaks of the PDFs.  We attribute these offsets to large
(random) statistical errors.
%ZH: I changed the wording above.  We didn't really prove that these are statistical errors,
%    and not some systematic shift to low power when N_s is low.
%AP: OK

Interestingly, we need as few as $N_s=2$ simulations to recover the
right PDF
%ZH: what is ``location parameter'' ?
%location parameter 
%AP: location parameter usually indicated where the PDF peaks. This might be the mean, but not always. 
for the small--scale power spectrum.  Figure~\ref{means_ns} shows that
multiple independent $N$--body simulations $N_s\geq2$ are indeed
necessary for measuring the means of feature ensembles to an accuracy
corresponding to 10\% of the statistical error. The number of required
simulations $N_s$ depends on the feature type and ranges from a few ($N_s=1$
or 2) for the power spectrum at low multipoles ($\ell\lesssim 500$) to
$N_s\approx 30-50$ for the power spectrum at larger
multipoles ($\ell\gtrsim1000$) or peak counts above a high threshold
($\kappa_0\approx0.3$).
%ZH: I think this text above is a bit misleading.  Wny would we actually impose the ``0.1 sigma''
%    requirement?    It sounds too stringent.  If we instead imposed, say, 0.5 sigma, would
%    N_s=2 always be sufficient?    We should probably at least add this (delete if untrue):
%AP: OK, agreed
On the other hand, relaxing the required accuracy to 50\% of the statistical
error, we find $N_s=2$ to be always sufficient.

Figure \ref{ps_var} shows the variance of convergence power spectrum
computed from different ensembles, in units of the Gaussian
expectation. We find that, even with $N_s=1$, we are able to recover
the known result that non--Gaussian structures increase the variance
significantly on small scales (see \citep{Sato12,TakadaSpergel14} for reference). 
%ZH: We could add more references here, e.g. the earlier Sato et al. 2009 paper, and the Takada and Spergel
%    papers looking at the non-Gaussian variance of the pow.spectrum.. 
%AP: Done: that is, in fact, the Sato 2009 paper, I mis-labelled it with Sato12 in the bibtex
Our results are in fact in excellent agreement with \citep{Sato12}, which used $N_s=400$ independent N-body
simulations.  This result is highly encouraging, suggesting that
individual $N$--body runs can be recycled repeatedly.  However, it is not
sufficient by itself to conclude that $N_s$ does not impact the
parameter inferences, since these depend on the cross band
covariances.

Figure \ref{curvingnb} investigates the parameter errors. This figure
shows that error degradation estimates truncated at order $O(1/N_r)$
are too optimistic when the number of simulations $N_r$ used to
measure the covariance is only a factor of few larger than the
dimension of the feature space $N_b$. In these cases effects coming
the next--to--leading orders $O(1/N_r^2)$ become non--negligible on
constraint degradation. In particular, we find that already for
$N_b=30$ and $N_r\sim100$, the error degradation estimates to the next
leading order, $O(1/N_r^2)$, remain too optimistic. Accurate analytic
estimates in this regime would require at least terms of order
$O(1/N_r^3)$, which come from higher--than--quartic $\bbh{\Psi}$
fluctations.

In Figure \ref{wvar_nr}, we examine how the degradation in the $w$
constraint depends on the number of simulations used to estimate the
covariance, in the limit of large $N_r$. We find excellent agreement
with the expected scaling (eq.~\ref{ourscaling}) up to values
comparable to $N_r\sim 10^5$. We therefore conclude that a single
$N$--body simulation is sufficient to construct an ensemble of
$O(10^5)$ mutually independent convergence power spectra.

Figure \ref{wvar_ns} shows how the ``true'' $w$ constraint (in the
limit $N_r\rightarrow \infty$; or equivalently the $w$ constraint with
the known $N_r$ dependence factored out),
%ZH: see my paraphrasing above.  I wanted to clearly distinguish the point
%    of this paragraph from the previous paragraph.
%AP: OK
depends on $N_s$. We find that, in the range $N_s\in[1,200]$ the
inferred $w$--variance $\sigma_{w,\infty}^2$ fluctuates stochastically
only by 1\%, and does not show any trend with $N_s$.

Finally, we found that when we estimate the data covariance $\bb{C}$
from the same simulation set used to measure $\bbh{\Psi}$, the
effective dimensionality $D$ decreases with increasing $N_b$ in the
case where the $\bbh{\Psi}$ bias is not corrected
(eq.~\ref{mockscalinguncorrected}).  This $N_b$-dependence disappears
when the bias is corrected (eq.~\ref{mockscalingcorrected}).  This
fact that should be taken into consideration when forecasting
parameter errors purely from simulations, as the errors will otherwise
be underestimated. A similar conclusion was reached by \citep{Hartlap07}.
%ZH: reworded above.  OK?  Also, I am not sure if this point is new.
%    Didn't Hartlap (and others) make this point? If so, we should cite them.
%    Also, it would make the point more relevant if we could cite some 
%    papers (including our own) which neglected this correction in the past
%    and therefore under-forecasted the errors.
%AP: yes, Hartlap et al. reached a similar conclusion. I think we are saying something more though: even if the bias in the Psi estimator is corrected, if the same simulation set for Psi and C is used, parameter errors are still underestimated because of the explanation we give in section II-B. I don't think this point appears in Hartlap et al.

%%%%%%%%%%%%%%%%%% CONCLUSION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}

In this work, we have examined the effect of forecasting cosmological
constraints based on shear ensembles generated from a finite number of
$N$--body simulations.  Our main results can be summarized as follows:
\vspace{0.4\baselineskip}
\begin{itemize}

\item When the feature covariance matrix is measured from simulations,
  parameter constraints are degraded. This degradation is appreciably
  larger than the $O(1/N_r)$ computed by \citep{DodelsonSchneider13}
  when the number of realizations $N_r$ is only a factor of few larger
  than the feature vector size $N_b$.
\item We can recycle a single $N$--body simulation to produce an
  ensemble of $O(10^5)$ independent maps. The mean feature measured
  from a shear ensemble, though, could be inaccurate if only one
  $N$--body simulation is used.
\item As few as one or two independent $N$--body simulations are
  sufficient to forecast $w$ errorbars to 1\% accuracy, provided that
  a sufficiently large number $N_r$ of realizations are used to
  measure feature covariances.  In particular, provided that biases
  in the inverse covariance are corrected, percent--level forecasts
  require $N_r\gsim 100 (N_b-N_p)$ realizations.
%ZH: what is sufficiently large here?  I presume a ``few hundred'' 
%    (or $>100N_p$, based on eq. 13).  I added that.  Please correct if 
%    this is not OK.
%AP: OK, the quantity to compare to is Nb-Np though. If Nb=Np there is no degradation at every order
\item Depending on the feature type used to constrain cosmology, a
  larger number of $N$--body simulations might be needed to measure
  accurate ensemble means to an accuracy corresponding 10\% of the statistical error. If this accuracy requirement is relaxed to 50\% of the statistical error, we find that as low as $N_s=2$ simulations are needed, for the feature types we consider in this work. 
%ZH:  So I am not sure about this last conclusion.   
%     What exactly do you have in mind?   Is it Figure 2, which shows
%     some deviations above the horizontal dashed line, for N_s up to ~50 ?
%     But this is based on the super-stringent requirement that the bias
%     is <0.1*sigma.
%     The same figure seems to show that N_s=2 simulation is in fact
%     sufficient to measure the means to within 0.5*sigma, as well.
%     If you agree, we should revise this bullet to state this more
%     positive conclusions!
%AP: Agreed, I rephrased the bullet point
\end{itemize}
%
Future extensions of this work should involve extending our analysis
to a larger set of cosmological parameters, and to more general
feature spaces, such as the ones that characterize non--Gaussian
statistics (e.g. including higher moments of the $\kappa$ field,
Minkowski Functionals, and higher-order $\kappa$ correlators). While
our results are highly encouraging, and suggest that a single N-body
simulation can be recycled repeatedly, to produce as many as $10^5$
independent shear power spectra or peak count histograms. In order to
scale our results to large future surveys, such as LSST, it will be
necessary to determine if our findinds hold when challenged by larger
and higher-resolution $N$--body simulations \citep{Qcontinuum}.

%%%%%%%%%%%%%%%%%%%%%%%%%% ACKNOWLEDGMENTS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Acknowledgements}
We thank Lam Hui for useful discussions.  The simulations in this work
were performed at the NSF XSEDE facility, supported by grant number
ACI-1053575, and at the New York Center for Computational Sciences, a
cooperative effort between Brookhaven National Laboratory and Stony
Brook University, supported in part by the State of New York. This
work was supported in part by the U.S. Department of Energy under
Contract Nos. DE-AC02-98CH10886 and DE-SC0012704, and by the NSF Grant
No. AST-1210877 (to Z.H.) and by the Research Opportunities and
Approaches to Data Science (ROADS) program at the Institute for Data
Sciences and Engineering at Columbia University (to Z.H.).

\bibliography{ref}

%%%%%%%%%%%%%%%%%%%%%%%%%% APPENDIX %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%ZH: I did not carefully re-read the appendices.  I guess that is OK.
%    One cosmetic comment:  when you refer to equations, you need to say
%    ``as in eq. (xx)''  rather than  ``as in (xx)''.
%AP: done   
\section*{Appendix A: cubic and quartic covariance fluctuations}
\label{appendixA}

The goal of this appendix is to give a derivation of eq. (\ref{quarticdegradation}). When the simulated feature vector $\bbh{d}_r$ is drawn from a Gaussian distribution, the covariance estimator $\bbh{C}$ follows the Wishart distribution, and its inverse $\bbh{\Psi}$ follows the inverse Wishart distribution. For the analytical expressions of these probability distributions we remand the reader to \citep{Taylor12}. Computing expectation values of eq. (\ref{estimatorcovariance}) over the inverse Wishart distribution is not possible analytically, and a perturbative expansion is necessary. Writing $\bbh{\Psi}=\bb{\Psi}+\delta\bbh{\Psi}$, we can expand eq. (\ref{quarticdegradation}) in powers of $\delta\bbh{\Psi}$. The expectation values of each term in this perturbative expansion can be calculated in term of the moments of the inverse Wishart distribution. \citep{MasumotoWishart} provide a general framework to compute these moments, and give exact expressions for moments up to quartic order. First, let us expand the inverse of the Fisher matrix estimator (\ref{estimatorfisher}) in powers of $\delta\bbh{\Psi}$. The $n$--th order of this expansion will be

\begin{equation}
\label{nthterm}
\delta\bbh{F}^{-1}_{(n)} = (-1)^{n}(\bb{F}^{-1}\delta\bbh{F})^n\bb{F}^{-1}
\end{equation}
%
with
\begin{equation}
\delta\bbh{F} = \bb{d}_0'^T\delta\bbh{\Psi}\bb{d}_0'
\end{equation}
%
Using eq. (\ref{nthterm}), we can expand eq. (\ref{estimatorcovariance}) to an arbitrary order in $\delta\bbh{\Psi}$, take the expectation values of the fluctuations over the inverse Wishart distribution, and finally get to eq. (\ref{quarticdegradation}). We use the notation $\nu=N_r-1$ and $\gamma=(\nu-N_b-1)/2$. We indicate with capital letters pairs of matrix indices, for example $I=(i_1,i_2)$, where $i_a=1..N_b$. The main results we take out from \citep{MasumotoWishart} regarding the first four moments are (up to order $O(1/\nu^2)$)

\begin{widetext}

\begin{equation}
\label{firstmoment}
\langle\h{\Psi}_I\rangle = \frac{\nu}{2\gamma}\Psi_I
\end{equation}

\begin{equation}
\label{secondmoment}
\langle\delta\h{\Psi}_I\delta\h{\Psi}_J\rangle = \frac{\nu^2\Psi_I\Psi_J + \nu^2\gamma\Psi_{[I}\Psi_{J]}}{4\gamma^2(\gamma-1)(2\gamma+1)}
\end{equation}

\begin{equation}
\label{thirdmoment}
\langle\delta\h{\Psi}_I\delta\h{\Psi}_J\delta\h{\Psi}_K\rangle = \frac{\nu^3\Psi_{[I}\Psi_J\Psi_{K]}}{8\gamma(\gamma-1)(\gamma-2)(\gamma+1)(2\gamma+1)}
\end{equation}

\begin{equation}
\label{fourthmoment}
\langle\delta\h{\Psi}_I\delta\h{\Psi}_J\delta\h{\Psi}_K\delta\h{\Psi}_L\rangle = \frac{\nu^4(2\gamma^2-5\gamma+9)\Psi_{[I}\Psi_{J]}\Psi_{[K}\Psi_{L]}}{16\gamma(\gamma-1)(\gamma-2)(\gamma-3)(2\gamma-1)(\gamma+1)(2\gamma+1)(2\gamma+3)}
\end{equation}

\end{widetext}
%
Where the square bracket notation is a shorthand for a symmetrization over pair of indices: for example
\begin{equation}
\Psi_{[I}\Psi_{J]} = \Psi_{i_1j_1}\Psi_{i_2j_2} + \Psi_{i_1j_2}\Psi_{i_2j_1}
\end{equation}
%
Eq. (\ref{firstmoment}) expresses the bias in the $\bbh{\Psi}$ estimator that already appears in the literature \citep{Hartlap07}. If we want to use the bias corrected $\bbh{\Psi}$ estimator (that is required in a perturbative expansion of eq. (\ref{estimatorcovariance})), we need to apply an additional factor of $(2\gamma/\nu)^n$ to eqs. (\ref{firstmoment}--\ref{fourthmoment}), where $n$ is the order of the moment we are applying the correction to. If we limit ourselves to computing the expectation value of eq. (\ref{estimatorcovariance}) up to order $O(1/\nu^2)$, we do not need to worry about this correction for eqs. (\ref{thirdmoment}--\ref{fourthmoment}), as the dominant term here is already $O(1/\nu^2)$. The next step is expanding eq. (\ref{estimatorcovariance}) in powers of $\delta\bbh{\Psi}$ up to fourth order: this is easily done

\begin{widetext}
\begin{equation}
\label{fullpertexpansion}
\h{\Sigma}_\bb{p} = \left(\bb{F}^{-1}+\sum_{n=1}^4\delta\bbh{F}^{-1}_{(n)}\right)\bb{d}'^T_0(\bb{\Psi}+\delta\bbh{\Psi})\bb{C}(\bb{\Psi}+\delta\bbh{\Psi})\bb{d}'_0\left(\bb{F}^{-1}+\sum_{n=1}^4\delta\bbh{F}^{-1}_{(n)}\right)
\end{equation}
\end{widetext}
%
Carrying out the calculations is simpler than it looks: because of the structure of eq. (\ref{fullpertexpansion}), each term in the expansion is going to be proportional to $\Sigma_\bb{p}f_a(N_b,N_p)/N_r^a$, where $f_a(N_b,N_p)$ is polynomial in $N_b,N_p$. Terms proportional to $N_b$ arise from index contractions of type $\mathrm{tr}(\bb{\Psi C})=N_b$, which come from symmetrization terms of type $\Psi_{[I}\Psi_{J]}$. Symmetrization factors of type $\Psi_{[I}\Psi_J\Psi_{K]}$ or $\Psi_{[I}\Psi_{J]}\Psi_{[K}\Psi_{L]}$, on the other hand, give rise to contractions of type $\mathrm{tr}(\bb{\Psi C})\mathrm{tr}(\bb{\bb{F}\bb{F}^{-1}})=N_bN_p$. Moreover, we know that, at every order $O(1/N_r^a)$, $f_a(N_b,N_p)$ has to be proportional to $N_b-N_p$ because it must vanish when $N_b=N_p$. The reason for this is that, when the feature derivative matrix $\bb{d}'_0$ is square and invertible (which it should be in absence of degeneracies), then eq. (\ref{estimatorcovariance}) reduces to

\begin{equation}
\h{\Sigma}_\bb{p} = (\bb{d}'_0)^{-1}\bb{C}(\bb{d}'^T_0)^{-1}
\end{equation} 
%
And every trace of the noise is gone, hence powers of $1/N_r^a$ must not appear at any order if $N_b=N_p$. Armed with the knowledge of the above considerations, we can compute the expectation value of eq. (\ref{fullpertexpansion}) at second, third and fourth order in $\delta\bbh{\Psi}$, keeping the terms that are at most $O(1/\nu^2)=O(1/N_r^2)$. When the combinatorial factors that arise from the expansion of eq. (\ref{fullpertexpansion}) are properly computed and the expectation values over the inverse Wishart distribution are taken according to eqs. (\ref{firstmoment}--\ref{fourthmoment}), the results should look like

\begin{widetext}
\begin{equation}
\label{expansionbreakdown}
\begin{cases}
\begin{displaystyle}
(\delta\bbh{\Psi})^2 \rightarrow \Sigma_\bb{p}\frac{\gamma(N_b-N_p)}{(\gamma-1)(2\gamma+1)} = \Sigma_\bb{p}\left[\frac{N_b-N_p}{N_r}+\frac{(N_b-N_p)(N_b+3)}{N_r^2}\right]
\end{displaystyle} \\ \\

\begin{displaystyle}
(\delta\bbh{\Psi})^3 \rightarrow -4\Sigma_\bb{p}\frac{(N_b-N_p)(1+N_p)}{N_r^2}
\end{displaystyle} \\ \\

\begin{displaystyle}
(\delta\bbh{\Psi})^4 \rightarrow 3\Sigma_\bb{p}\frac{(N_b-N_p)(1+N_p)}{N_r^2}
\end{displaystyle}

\end{cases}
\end{equation}
\end{widetext}
%
When the results from eq. (\ref{expansionbreakdown}) are summed, eq. (\ref{quarticdegradation}) immediately follows.

\clearpage
\newpage
\section*{Appendix B: negative effective dimensionality}
\label{appendixB}

The goal of this appendix is giving a justification on why the effective dimensionality $D$ that appears in eq. (\ref{ourscaling}) can be negative in some cases. When we use the same simulation set to estimate $\bb{C},\bbh{\Psi}$, eq. (\ref{estimatorcovariance}) reduces to the inverse Fisher estimator $\h{\Sigma}_\bb{p}=\bbh{F}^{-1}$. At second order in the $\Psi$ fluctuations this becomes

\begin{equation}
\h{\Sigma}_\bb{p} = \bb{F}^{-1} + \bb{F}^{-1}\left(- \delta{\bbh{F}}+ \delta{\bbh{F}}\bb{F}^{-1}\delta{\bbh{F}}\right)\bb{F}^{-1}  
\end{equation}  
%
If the biased estimator for $\bbh{\Psi}$ is used, we can use eqs. (\ref{firstmoment}--\ref{secondmoment}) at order $O(1/\nu)$ to compute 
\begin{widetext}
\begin{equation}
\label{negativeD}
\langle\h{\Sigma}_\bb{p}\rangle = \Sigma_\bb{p}\left(1-\frac{N_b+1}{N_r}+\frac{1+N_p}{N_r}\right) = \Sigma_\bb{p}\left(1+\frac{N_p-N_b}{N_r}\right)
\end{equation}
\end{widetext}
%
And we immediately see that the coefficient of $1/N_r$ is negative, because $N_b>N_p$. This is the result shown in eq. (\ref{mockscalinguncorrected}). If the bias correction for $\bbh{\Psi}$ is applied, the first order terms $\delta{\bbh{F}}$ average to 0, and we are left with only the last term in the sum (\ref{negativeD}), which immediately yields eq. (\ref{mockscalingcorrected}).

\label{lastpage}
\end{document}